---
title: "Hands-on_Ex01"
author: "Cathy Chu"
date-modified: "last-modified" 
execute:
  echo: true 
  eval: true 
  warning: false 
  freeze: true  
---

## A Layered Grammar of Graphics: ggplot2 methods

### 1.2 Getting started

#### **1.2.1 Installing and loading the required libraries**

pacman package has been installed in the workshop, using the following code.

```{r}
pacman::p_load(tidyverse)
```

#### **1.2.2 Importing data**

*read_csv* is a function of **readr** package, which is from the tidyverse package.

The dataset Exam_data.csv, containing exam grades of P3 students from a Singapore local school is imported into R environment. The code chunk below was used to import the dataset.

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

|            |           |            |          |           |             |             |
|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
| **ID**     | **CLASS** | **GENDER** | **RACE** | **MATHS** | **ENGLISH** | **SCIENCE** |
| Student321 | 3I        | Male       | Malay    | 21        | 9           | 15          |
| Student305 | 3I        | Female     | Malay    | 24        | 22          | 1           |

: {.striped .hover}

------------------------------------------------------------------------

### **1.3 Introducing ggplot**

**ggplot2** is part of the tidyverse family specially designed for visual exploration and communication. It creates graphics based on The Grammar of Graphics.

#### ðŸ“ Cheatsheet for ggplot2 ([learn more...](https://ggplot2.tidyverse.org/#cheatsheet))

![](images/data-visualization-01.pdf){width="400" height="250"}
