---
title: "In-class_Ex07"
author: "Cathy C."
date-modified: "last-modified" 
execute:
  echo: true 
  eval: true 
  warning: false 
  freeze: true
editor: visual
---

## 1 Getting started

**Load and launch packages**

Use tool to install packages: **tsibble, feasts, fable and seasonal** first.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
pacman::p_load(tidyverse, tsibble, feasts, fable, seasonal)
```

## 2 Import the data

### 2.1 Import csv data

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ts_data <- read_csv("data/visitor_arrivals_by_air.csv")
head(ts_data)
```

In the code chunk below, `dmy()` of **lubridate** package is used to convert data type of Month-Year field from Character to Date.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ts_data$`Month-Year` <- dmy(
  ts_data$`Month-Year`) #to replace the column.
head(ts_data)
```

Now we see it's converted to **date** data type.

### 2.2 **Conventional base `ts` object versus `tibble` object**

#### 2.2.1 `ts_data` is a tibble object

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ts_data
```

We can also use class() to check.

#### 2.2.2 `ts_data_ts` to be a ts object

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ts_data_ts <- ts(ts_data)       
head(ts_data_ts)
```

Check how the conversion goes.

```{r}

class(ts_data)
class(ts_data_ts)
```

### 2.3 **Converting `tibble` object to `tsibble` object**

A **tsibble** (or tbl_ts) is a data- and model-oriented object. Compared to the conventional time series objects in R, for example ts, zoo, and xts, the tsibble preserves time indices as the essential data column and makes heterogeneous data structures possible. Beyond the tibble-like representation, key comprised of single or multiple variables is introduced to uniquely identify observational units over time (index).

The code chunk below converting ts_data from tibble object into tsibble object by using [`as_tsibble()`](https://tsibble.tidyverts.org/reference/as-tsibble.html) of **tsibble** R package.

```{r}

ts_tsibble <- ts_data %>%
  mutate(Month = yearmonth(`Month-Year`)) %>%
  as_tsibble(index = `Month`)
```

::::: goals
::: goals-header
Learning from the code
:::

::: goals-container
mutate() of dplyr package is used to derive a new field by transforming the data values in Month-Year field into month-year format. The transformation is performed by using yearmonth() of tsibble package. + as_tsibble() is used to convert the tibble data frame into tsibble data frame.
:::
:::::

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ts_tsibble
```

## **3 Visualising Time-series Data**

In order to visualise the time-series data effectively, we need to organise the data frame from wide to long format by using [`pivot_longer()`](https://tidyr.tidyverse.org/reference/pivot_longer.html) of **tidyr** package.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ts_longer <- ts_data %>%
  pivot_longer(cols = c(2:34),
               names_to = "Country",
               values_to = "Arrivals")
```

### 3.1 **Visualising single time-series: ggplot2 methods**

```{r}
#| code-fold: true
#| code-summary: "Show the code"
ts_longer %>%
  filter(Country == "Vietnam") %>%
  ggplot(aes(x = `Month-Year`,
             y = Arrivals)) +
  geom_line(size = 0.5, color = "#3AA6B9")
```

::::: goals
::: goals-header
Learning from the code
:::

::: goals-container
-   [`filter()`](https://dplyr.tidyverse.org/reference/filter.html) of [**dplyr**](https://dplyr.tidyverse.org/) package is used to select records belong to Vietnam.

-   [`geom_line()`](https://ggplot2.tidyverse.org/reference/geom_path.html) of [**ggplot2**](https://ggplot2.tidyverse.org/) package is used to plot the time-series line graph.
:::
:::::

### 3.2 **Plotting multiple time-series data with ggplot2 methods**

To plot multiple countries

```{r}
#| fig-width: 9
#| fig-height: 10
#| code-fold: true
#| code-summary: "Show the code"
  
ggplot(data = ts_longer, 
       aes(x = `Month-Year`, 
           y = Arrivals,
           color = Country))+
  geom_line(size = 0.5) +
  theme(legend.position = "bottom", 
        legend.box.spacing = unit(0.5, "cm"))
```

In order to provide effective comparison, [`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html) of **ggplot2** package is used to create small multiple line graph also known as trellis plot.

```{r}
#| fig-width: 7
#| fig-height: 15
ggplot(data = ts_longer, 
       aes(x = `Month-Year`, 
           y = Arrivals))+
  geom_line(size = 0.5) +
  facet_wrap(~ Country,
             ncol = 3,
             scales = "free_y") + #y is not fixed axis.
  theme_bw()
```

## 4 Visual Analysis of Time-series Data

```{r}
tsibble_longer <- ts_tsibble %>%
  pivot_longer(cols = c(2:34),
               names_to = "Country",
               values_to = "Arrivals")
```

### 4.1 **Visual Analysis of Seasonality with Seasonal Plot**

A **seasonal plot** is similar to a time plot except that the data are plotted against the individual **seasons** in which the data were observed.

A season plot is created by using [`gg_season()`](https://feasts.tidyverts.org/reference/gg_season.html) of **feasts** package. Below assume season is MONTH.

```{r}
tsibble_longer %>%
  filter(Country == "Italy" |
         Country == "Vietnam" |
         Country == "United Kingdom" |
         Country == "Germany") %>% 
  gg_season(Arrivals)
```

### 4.2 **Visual Analysis of Seasonality with Cycle Plot**

A **cycle plot** shows how a trend or cycle changes over time. We can use them to see seasonal patterns. Typically, a cycle plot shows a measure on the Y-axis and then shows a time period (such as months or seasons) along the X-axis. For each time period, there is a trend line across a number of years.

We can start with the plot below, before digging deeper. Note that both lines reveal clear sign of seasonal patterns but not the trend.

```{r}
tsibble_longer %>%
  filter(Country == "Vietnam" |
         Country == "Italy") %>% 
  autoplot(Arrivals) + 
  facet_grid(Country ~ ., scales = "free_y")
```

*\<feast package\>* In the code chunk below, cycle plots using [`gg_subseries()`](https://feasts.tidyverts.org/reference/gg_subseries.html) of feasts package are created. Notice that the cycle plots show not only seasonal patterns but also trend, by **month**.

```{r}
#| fig-width: 6
tsibble_longer %>%
  filter(Country == "Vietnam" |
         Country == "Italy") %>% 
  gg_subseries(Arrivals) 
```

-   increasing trend.

-   rate high

-   July and August are higher than others.

-   Italy: most months are very low, except Aug.

## 5 Time series decomposition

**Trend, seasonal and error**

Time series decomposition allows us to isolate structural components such as trend and seasonality from the time-series data.

### 5.1 **Single time series decomposition**

In **feasts** package, time series decomposition is supported by [`ACF()`](https://feasts.tidyverts.org/reference/ACF.html), `PACF()`, `CCF()`, [`feat_acf()`](https://feasts.tidyverts.org/reference/feat_acf.html), and [`feat_pacf()`](https://feasts.tidyverts.org/reference/feat_acf.html). The output can then be plotted by using `autoplot()` of **feasts** package.

In the code chunk below, `ACF()` of **feasts** package is used to plot the ACF curve of visitor arrival from Vietnam.

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  ACF(Arrivals) %>% 
  autoplot()
```

In the code chunk below, `PACF()` of **feasts** package is used to plot the Partial ACF curve of visitor arrival from Vietnam.

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  PACF(Arrivals) %>% 
  autoplot()
```

### 5.2 **Multiple time-series decomposition**

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam" |
         `Country` == "Italy" |
         `Country` == "United Kingdom" |
         `Country` == "China") %>%
  ACF(Arrivals) %>%
  autoplot()
```

::::: goals
::: goals-header
Observation
:::

::: goals-container
`ACF()`: compare with the previous (relation with the previous)
`autoplot()` gives the plotting.
-   China and Vietnam are similar, but not the same. China is 6 months period, while VN is 12 months.
-   blues lines are confidence intervals to check whether they are statistically significant. China is more significant then VN. The rest are weak. UK - trend is not significant, but still have seasonal.
:::
:::::

On the other hand, code chunk below is used to prepare a trellis plot of **PACFs** for visitor arrivals from Vietnam, Italy, United Kingdom and China.

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam" |
         `Country` == "Italy" |
         `Country` == "United Kingdom" |
         `Country` == "China") %>%
  PACF(Arrivals) %>%
  autoplot()
```
