---
title: "Hands-on_Ex01"
author: "Cathy Chu"
date-modified: "last-modified" 
execute:
  echo: true 
  eval: true 
  warning: false 
  freeze: true  
---

# **Choropleth Mapping with R**

## 1 Overview

Choropleth mapping involves the symbolisation of enumeration units, such as countries, provinces, states, counties or census units, using area patterns or graduated colors.

For example, a social scientist may need to use a choropleth map to portray the spatial distribution of aged population of Singapore by Master Plan 2014 Subzone Boundary.

We will learn how to plot functional and truthful choropleth maps by using an R package called [**tmap**](https://cran.r-project.org/web/packages/tmap/) package.

::::: goals
::: goals-header
Tips
:::

::: goals-container
Read each functional description before use.
:::
:::::

## 2 Getting Started

In this hands-on exercise, the key R package use is [tmap](https://cran.r-project.org/web/packages/tmap/) package in R. Besides **tmap** package, four other R packages will be used. They are:

-   [**readr**](https://readr.tidyverse.org/) for importing delimited text file

-   [**tidyr**](https://tidyr.tidyverse.org/) for tidying data

-   [**dplyr**](https://dplyr.tidyverse.org/) for wrangling data

-   [sf](https://cran.r-project.org/web/packages/sf/index.html) for handling geospatial data

Among the four packages, **readr, tidyr** and **dplyr** are part of **tidyverse** package.

The code chunk below will be used to install and load these packages in RStudio.

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

::::: goals
::: goals-header
Tips
:::

::: goals-container
Only need to install **tidyverse** instead of **readr, tidyr** and **dplyr** indivisually.
:::
:::::

## 3 Import Data

### 3.1 Data

Two data set will be used to create the choropleth map. They are:

-   Master Plan 2014 Subzone Boundary (Web) (i.e. `MP14_SUBZONE_WEB_PL`) in ESRI shapefile format. It can be downloaded at [data.gov.sg](https://data.gov.sg/) This is a geospatial data. It consists of the geographical boundary of Singapore at the planning subzone level. The data is based on URA Master Plan 2014.

-   Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2011-2020 in csv format (i.e. `respopagesextod2011to2020.csv`). This is an aspatial data fie. It can be downloaded at [Department of Statistics, Singapore](https://www.singstat.gov.sg/). Although it does not contain any coordinates values, its PA and SZ fields can be used as unique identifiers to geocode to `MP14_SUBZONE_WEB_PL` shapefile.

### 3.2 Import geospatial data to R

The code chunk below uses the *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL` shapefile into R as a simple feature data frame called `mpsz`.

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

To examine the data:

```{r}
mpsz
```

::::: goals
::: goals-header
Why only 10 records are displayed?
:::

::: goals-container
This is a default behavior in R, particularly for spatial data frames and sf (simple features) objects.

A few ways to view more rows:

-   Use `print(mpsz, n = 20)` to show more rows (20 in this example)
-   Use `View(mpsz)` to open the data in R's data viewer window
-   Use `head(mpsz, 30)` to see the first 30 rows
-   Use `str(mpsz)` to see the structure of the entire object
-   Use `summary(mpsz)` to get a statistical summary
:::
:::::

### 3.3 Import attribute data to R

Next, we will import *respopagsex2011to2020.csv* file into RStudio and save the file into an R dataframe called *popagsex*.

The task will be performed by using *`read_csv()`* function of **readr** package as shown in the code chunk below.

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

### 3.4 Data prep

Before a thematic map can be prepared, it is required to prepare a **data table with year 2020 values**. The data table should include the variables P**A, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.**

-   YOUNG: age group 0 to 4 until age groyup 20 to 24
-   ECONOMY ACTIVE: age group 25-29 until age group 60-64
-   AGED: age group 65 and above
-   TOTAL: all age group
-   DEPENDENCY: the ratio between young and aged against economy active group

#### 3.4.1 Data wrangling

The following data wrangling and transformation functions will be used:

-   *pivot_wider()* of **tidyr** package
-   *mutate(), filter(), group_by()* and *select()* of **dplyr** package

```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

#### 3.4.2 Joining the attribute data and geospatial data

Before we can perform the georelational join, one extra step is required to convert the values in PA and SZ fields to **uppercase**. This is because the values of PA and SZ fields are made up of upper- and lowercase. On the other, hand the *SUBZONE_N* and *PLN_AREA_N* are in uppercase.

```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

Next, *left_join()* of **dplyr** is used to join the geographical data and attribute table using planning subzone name e.g. *SUBZONE_N* and *SZ* as the common identifier.

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

Thing to learn from the code chunk above:

*left_join()* of **dplyr** package is used with `mpsz` simple feature data frame as the left data table is to ensure that the output will be a simple features data frame.

```{r}
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## 4 **Choropleth Mapping Geospatial Data Using *tmap***

Two approaches can be used to prepare thematic map using tmap, they are: - Plotting a thematic map quickly by using qtm(). - Plotting highly customisable thematic map by using tmap elements.

### 4.1 \*\*Plotting a choropleth map quickly by using \*qtm()\*\*

The easiest and quickest to draw a choropleth map using **tmap** is using *qtm()*. It is concise and provides a good default visualisation in many cases.

The code chunk below will draw a cartographic standard choropleth map as shown below.

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

```{r}
tmap_mode("view")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

::::: goals
::: goals-header
Learn from code chunk
:::

::: goals-container
-   *tmap_mode()* with “plot” option is used to produce a static map. For interactive mode, “view” option should be used.

-   *fill* argument is used to map the attribute (i.e. DEPENDENCY)
:::
:::::

### 4.2 **Creating a choropleth map by using *tmap*’s elements**

Despite its usefulness of drawing a choropleth map quickly and easily, the disadvantge of *qtm()* is that it makes aesthetics of individual layers harder to control. To draw a high quality cartographic choropleth map as shown in the figure below, **tmap**’s drawing elements should be used.

```{r}
#| code-fold: true
#| code-summary: "Show the code"


tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

In the following sub-section, we will learn the **tmap** functions that used to plot these elements.

#### 4.2.1 Drawing a base map

The basic building block of **tmap** is *tm_shape()* followed by one or more layer elemments such as *tm_fill()* and *tm_polygons()*.

In the code chunk below, *tm_shape()* is used to define the input data (i.e *mpsz_pop2020*) and *tm_polygons()* is used to draw the planning subzone polygons

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

#### 4.2.2 Drawing a choropleth map using *tm_polygons()*

To draw a choropleth map showing the geographical distribution of a selected variable by planning subzone, we just need to **assign the target variable** such as *Dependency* to *tm_polygons()*.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

::::: goals
::: goals-header
Learn from *tm_polygons()*
:::

::: goals-container
-   The default interval binning used to draw the choropleth map is called “pretty”. A detailed discussion of the data classification methods supported by **tmap** will be provided in sub-section 4.3.

-   The default colour scheme used is `YlOrRd` of ColorBrewer. You will learn more about the color scheme in sub-section 4.4.

-   By default, Missing value will be shaded in grey.
:::
:::::

#### 4.2.3 Drawing a choropleth map using *tm_fill()* and *tm_border()*

*tm_polygons() = tm_fill()* + *tm_border()*

*tm_polygons()* is a wraper of *tm_fill()* and *tm_border()*. *tm_fill()* shades the polygons by using the default colour scheme and *tm_borders()* adds the borders of the shapefile onto the choropleth map.

The code chunk below draws a choropleth map by using *tm_fill()* alone.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

Notice that the planning subzones are **shared** according to the respective dependecy values

To add the boundary of the planning subzones, *tm_borders* will be used as shown in the code chunk below.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  fill_alpha = 1)
```

### 4.3 **Data classification methods of tmap**

### 4.4 Color scheme

### 4.5 Map layouts

### 4.6 Drawing small multiple choropleth maps

### 4.7 Mapping spatial object meeting a selection criterion

## Reference

### 5.5 Plotting the cycle plot

## 6 Plotting Slopegraph

### 6.1 Data import

### 6.2 Plotting the slopegraph
