---
title: "Hands-on_Ex08_3"
author: "Cathy Chu"
date-modified: "last-modified" 
execute:
  echo: true 
  eval: true 
  warning: false 
  freeze: true  
---

# **Analytical Mapping**

## 1 Overview

### 1.1 **Objectives**

-   Importing geospatial data in rds format into R environment.

-   Creating cartographic quality choropleth maps by using appropriate tmap functions.

-   Creating rate map

-   Creating percentile map

-   Creating boxmap

## 2 Getting Started

### 2.1 Install and load packages

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

### 2.1 Import data to R

For the purpose of this hands-on exercise, a prepared data set called *NGA_wp.rds* will be used. The data set is a polygon feature data.frame providing information on water point of Nigeria at the LGA level.

```{r}
nga_wp <- read_rds("data/rds/NGA_wp.rds")
```

## 3 Basic choropleth mapping

### 3.1 **Visualising distribution of non-functional water point**

## 4 **Choropleth Map for Rates**

### 4.4 **Deriving Proportion of Functional Water Points and Non-Functional Water Points**

### 4.5 **Plotting map of rate**

## 5 **Extreme Value Maps**

### 5.1 **Percentile Map**

#### **5.1.1 Data Preparation**

#### **5.1.2 Why writing functions?**

#### **5.1.3 Creating the get.var function**

#### **5.1.4 A percentile mapping function**

### 5.2 **Box map**

#### **5.2.1 Creating the boxbreaks function**

#### **5.2.2 Creating the get.var function**

#### **5.2.3 Test drive the newly created function**

#### **5.2.4 Boxmap function**

## Reference

### 
