---
title: "Hands-on_Ex01"
author: "Cathy Chu"
date-modified: "last-modified" 
execute:
  echo: true 
  eval: true 
  warning: false 
  freeze: true  
---

# **Visualising and Analysing Time-oriented Data**

## 1 Learning Objectives

In this chapter, we will learn to plot the following visualisations:

-   a calender heatmap (ggplot2 functions)
-   a cycle plot (ggplot2 functions)
-   a slopegraph
-   a horizon chart

## 2 Getting Started

To install and launch the following R packages:

**scales, viridis, lubridate, ggthemes, gridExtra, readxl, knitr, data.table and tidyverse.**

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes, gridExtra, readxl, knitr, 
               data.table, tidyverse, CGPfunctions, ggHoriPlot)
```

## 3 Plotting Calendar Heatmap

Within this section, we will do the following:

-   plot a calendar heatmap by using ggplot2 functions and extension

-   write function using R programming

-   derive specific date and time related field by using base **R** and **libridate** packages

-   perform data preparation task by using tidyr and dplyr packages

### 3.1 Data

*eventlog.csv* file will be used for this exercise. This data file consists of 199,999 rows of time-series cyber attach records by country.

::: panel.tabset
## Import data

To import *eventlog.csv* into R environment and call this data frame as ***attacks.***

```{r}
attacks <- read_csv(data/eventlog.csv)
```

## Exam data structure

## Data prep
:::

### 3.2 Build the calendar heatmaps

### 3.3 Build multiple calendar heatmaps

## 4 Plotting Cycle Plot

### 4.1 Data import

### 4.2 Deriving month and year fields

### 4.3 Extract the target country

### 4.4 Computing year average arrivals by month

### 4.5 Plotting the cycle plot

## 5 Plotting Slopegraph

### 5.1 Data import

### 5.2 Plotting the slopegraph
